---
title: "Project1"
author: "Jenny Norton"
date: "9/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The Dataset: UFO Sightings Around the World

The data for this project was obtained from https://www.kaggle.com/camnugent/ufo-sightings-around-the-world by Cam Nugent.

## Loading and Cleaning the Data

You can load the clean data here:

```{r}
source("CleanData.R")
```

I read the data into ufo_data and changed the date_documented column from col_character to col_date with the formatting to remove a character at the end.

I had four parsing failures in the length_of_encounter_seconds column and the latitude column but I decided not to fix it at this point.  

I tried parsing Date_time to a date and time format but there were 698 parsing failures do to the time being 24:00.  Using a str_replace command, I was able to search and replace the 24:00 with 00:00 and parse into the date and time format and add a new column called Date_time2.

I didn't like how they had the state/province column named so I did a rename command to change it from `state/province` to state_province.

I filtered the data by the Date_time2 column for incidents that happened in the 1970s and arranged by shape and saved it as ufo_data1970 in case I wanted to work with a smaller subset of the data.


## Overview

The data set contains 80,332 entries, dating back to 1949, with 11 original columns:

Date_time - standardized date and time of sighting
city - location of UFO sighting
state_province - the US state or Canadian province, appears blank for other locations
country - Country of UFO sighting
UFO_shape - a one word description of the "spacecraft"
length_of_encounter_seconds - standardized to seconds, length of the observation of the UFO
described_duration _of_encounter - raw description of the length of the encounter 
description - text description of the UFO encounter
date_documented - when was the UFO sighting reported
latitude - latitude
longitude - longitude

## Summary

The data set contains over 80,000 records of UFO sightings back to 1949.  The original file was accessed by Cam Nugent at https://github.com/planetsig/ufo-reports which was prepared by Sigmond Axel in 2014.  Sigmond received the data from NUFORC (National UFO Reporting Center) at www.nuforc.org and did a bunch of cleaning of the dates which were using two digit years.

At this point, I'm interested in the UFO_shape and the Date_time2 columns.  I might also use the state_province and the country columns after further review and cleaning and the length_of_encounter_seconds column.

## Exploration of the Data

What are the different UFO Shapes?

```{r}
ufo_data %>%
  count(UFO_shape) %>%
  arrange(desc(n))
```

What states or provinces have the most sightings?

```{r}
ufo_data %>%
  count(state_province) %>%
  arrange(desc(n))
```

I see there are quite a few records without a state or province listed so let's filter those out.

```{r}
ufo_data %>%
  filter(!is.na(state_province)) %>%
  count(state_province) %>%
  arrange(desc(n))
```

Looks like California has the most sightings.  I wonder where New Mexico shows up on this list?

```{r}
ufo_data %>%
  filter(!is.na(state_province), state_province == "nm") %>%
  count(state_province) %>%
  arrange(desc(n))
```

```{r}
ufo_data %>%
  filter(!is.na(state_province)) %>%
  ggplot() +
  geom_bar(aes(x = state_province)) +
  coord_cartesian(xlim = c(0, 10)) +
  labs(title = "Number of UFO Sightings by State or Province",
       x = "State or Province",
       y = "Sightings")
```

```{r}
ufo_data %>%
  filter(!is.na(UFO_shape), !is.na(state_province), state_province == "ca") %>%
ggplot() +
      geom_bar(aes(x = UFO_shape)) +
               coord_flip() +
  labs(title = "CA State UFO Data by Shape", x = "UFO Shape", y = "Number in CA")
```

```{r}
ufo_data %>%
  group_by(UFO_shape) %>%
  mutate(ufo_shape_total = n()) -> ufo_data
```

```{r}
ufo_data %>%
  summarize(state_province, ufo_shape_total)
```

What are the top 10 states with light sightings?
```{r}
ufo_data %>%
  filter(!is.na(state_province), UFO_shape == "light") %>%
  count(state_province) %>%
  arrange(desc(n)) %>%
  top_n(10)
```

How can I graph this?
```{r}
ufo_data %>%
  filter(UFO_shape == "light") %>%
  group_by(state_province) %>%
  filter(n() > 426) %>%
  ungroup() %>%
  ggplot() +
  geom_bar(aes(x = state_province), fill = "blue") +
  labs(title = "Top 10 States/Provinces with 'Light' Sightings",
       x = "State/Province")
```

I wonder why California has so many more light sightings than the other states?  The next highest shape is the triangle.  I wonder what state has the most sightings of the triangle?

```{r}
ufo_data %>%
  filter(!is.na(state_province), UFO_shape == "triangle") %>%
  count(state_province) %>%
  arrange(desc(n)) %>%
  top_n(10)
```

```{r}
ufo_data %>%
  filter(UFO_shape == "triangle") %>%
  group_by(state_province) %>%
  filter(n() > 238) %>%
  ungroup() %>%
  ggplot() +
  geom_bar(aes(x = state_province), fill = "green") +
  labs(title = "Top 10 States/Provinces with 'Triangle' Sightings",
       x = "State/Province")
```

California again which we should have gathered since California has the most overall sightings. 

## Future Work

I would like to keep working with my data and see if I can:

  * Determine if the UFO_shape and the length_of_encounter_seconds have any relationship.  
  * See the top 10 states with overall sightings. 
  * Calculate the length_of_encounter_seconds into minutes or maybe hours.
  * Calculate some averages or medians for the length_of_encounter.
  * Make some really cool graphs.
  
Jenny Norton


